---
title: "BST270 Individual Project"
author: "Ziqi Fu"
date: "2025-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'hold', echo = FALSE, warning = FALSE, message = FALSE)
```

## Dependencies
This notebook can be reproduced by installing the following R packages: 

- ggplot2
- dplyr
- readr
- tidyr
- magrittr
- knitr

```{r, message=FALSE, warning=FALSE,echo=TRUE}
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(magrittr)
library(knitr)
```

For packages that are not installed, use the command `install.packages('PACKAGE_NAME')` for installation.

## 1. Reproducing the table in article [Where Are America’s Librarians?](https://fivethirtyeight.com/features/where-are-americas-librarians/)

We first download the data file `librarians-by-msa.csv` from its [GitHub repo](https://github.com/fivethirtyeight/data/tree/master/librarians). A copy of this data file has also been stored locally in the `Data` directory. Detailed [codes](#codes-1) are available in the end of this file.

```{r librarian, echo=FALSE}
# Load and clean data
lib.data=read.csv("Data/librarians-by-msa.csv", header=T)  
names(lib.data)=tolower(names(lib.data))
lib.data$tot_emp=as.numeric(gsub("[$]|,", "", lib.data$tot_emp))
lib.data$emp_prse=as.numeric(gsub("[$]|,", "", lib.data$emp_prse))
lib.data$jobs_1000=as.double(lib.data$jobs_1000) %>% round(2)
lib.data = drop_na(lib.data)

# Build model and get coefficients
nl.model=nls(data=lib.data, formula=emp_prse~a*tot_emp^b, start=list(a=1, b=1))
a=coef(nl.model)[1]
b=coef(nl.model)[2] 

# Get lower and upper estimates
lib.data$mor=(a*(lib.data$tot_emp^b))*1 #SD=1
lib.data$HE=(lib.data$tot_emp*(1+(lib.data$mor/100))) %>% round(0)
lib.data$LE=(lib.data$tot_emp*(1-(lib.data$mor/100))) %>% round(0)

# select areas to generate a table
select.area = c('Owensboro, KY',
                 'Nassau-Suffolk, NY Metropolitan Division',
                 'Bethesda-Rockville-Frederick, MD Metropolitan Division',
                 'New Haven, CT',
                 'Haverhill-North Andover-Amesbury, MA-NH NECTA Division',
                 'Tallahassee, FL',
                 'Durham-Chapel Hill, NC',
                 'Ithaca, NY',
                 'Dover, DE',
                 'St. Joseph, MO-KS')

tb = lib.data[match(select.area,lib.data$area_name),c(2,5,3,9,8)] %>% as.data.frame()
colnames(tb) = c('METROPOLITAN AREA','JOB QUOTIENT','NUMBER OF LIBRARIANS','LOWER ESTIMATE','UPPER ESTIMATE')
kable(tb,row.names = F) # display the table
```

### Notes
The lower and upper estimates are computed using `SD=1`, and the original data does not contain information about `HAS A COLLEGE` for the selected areas. For simplicity, we will skip the last columns, since all the selected areas are indicated to have a college in the original table we try to replicate. All the numbers in this table match perfectly with the original table. 


## 2. Reproducing the first graph in article [The Dallas Shooting Was Among The Deadliest For Police In U.S. History](https://fivethirtyeight.com/features/the-dallas-shooting-was-among-the-deadliest-for-police-in-u-s-history/)

We first download the cleaned version of the data `clean_data.csv` from its [GitHub repo](https://github.com/fivethirtyeight/data/tree/master/police-deaths). This data file has been renamed as `police-death-clean.csv` and stored under the `/Data` directory. We are reproducing the first graph in the original article, and the detailed [codes](#codes-2) are attached below.

```{r police,echo=FALSE}
# Load and clean data
clean_data = read.csv('Data/police-death-clean.csv') %>% filter(canine == FALSE) 

# Major categories found in the data
large_categories = clean_data %>%
      group_by(year, cause_short) %>%
      summarize(count = n()) %>%
      data.frame() %>%
      filter(count >= 20) %>% # chose the category if # incidence >= 20
      select(cause_short) %>%
      unique()
cat_to_plot = c(large_categories$cause_short, "Gunfire (Accidental)") #also include accidental gunfire cases

plot_order = clean_data %>%
    mutate(cat = ifelse(cause_short %in% cat_to_plot, cause_short, 'other')) %>%
    group_by(cat) %>%
    summarize(count = n()) %>%
    data.frame() %>%
    arrange(desc(count)) %>%
    extract2(1)

# Move order to the end
plot_order = c(plot_order[! (plot_order == 'other')], 'other')
    
# Create data for plotting
data_for_plot = clean_data %>%
      mutate(cat = ifelse(cause_short %in% cat_to_plot, cause_short, 'other')) %>%
      group_by(year, cat) %>%
      summarize(count = n()) %>%
      data.frame() %>%
      spread(cat, count)
data_for_plot[is.na(data_for_plot)] = 0
data_for_plot = data_for_plot %>%
      gather(cat, count, -year) %>%
      mutate(cat = factor(cat, levels=plot_order)) %>%
      arrange(cat)


# generate the plot
ggplot(data_for_plot, aes(x=year, y=count, group=cat, order=cat)) +
      geom_area(aes(fill=cat),color='blue',size=0.05, position='stack') +
      labs(fill=NULL,title='On-duty police officer deaths in the U.S.',subtitle='By cause since 1791',x=NULL,y=NULL)+
      theme(legend.position = c(0.2,0.55),legend.background = element_rect(fill = "grey"))

```

### Notes
The graph is highly reproducible except for a few aesthetic differences. Here, we directly use the cleaned data set and avoid any further data wrangler needed to process the raw data. The raw data is also available on the article's GitHub page, and the authors uploaded the data cleaning script for reproducibility. 

\section{Codes} 
## Codes for reproducing table 1 {#codes-1}
```{r ref.label='librarian', echo=TRUE, eval=FALSE} 

```

## Codes for reproducing graph 1 {#codes-2}
```{r ref.label='police', echo=TRUE, eval=FALSE} 

```


